---
title: "Candy Counting and the Crowd"
author: "Stephen Kaluzny"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Challenge

Guess the number of pieces of pieces of candy in the jar.
This was a typically activity at the Seattle Office's Halloween Party.
There you could handle the physical jar of candy.
For this exercise, several photos of the jar were provided.

There is an idea that the Wisdom of the Crowd would provide a good
estimate of the count.
The story goes that guesses for a pig's weight at a state fair were highly
variable but the average of all the guesses was very close to the correct value.

The exercise was made available to three groups, the Seattle office (TIBCO),
my extended family (Family),
and the Seattle office alumni group (StatSoftFul).

```{r load_packages, echo=FALSE}
suppressPackageStartupMessages(library("here"))
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("ggplot2"))
```

## The Data

```{r read_data}
guesses <- read.csv(here::here("data", "candyjar_guesses.csv"))
```
There are `r nrow(guesses)` total observations (guesses),
`r sum(guesses$Group == "TIBCO")` from TIBCO,
`r sum(guesses$Group == "Family")` from my family,
and `r sum(guesses$Group == "StatSoftFul")` from Seattle office alumni.

Compute the true candy count based on the counts of the types of
candy pieces:
```{r true_count}
true_count <- read.csv(here::here("data", "candyjar_count.csv"))
Count <- true_count[, "Count", drop=TRUE]
names(Count) <- true_count[, "Type", drop=TRUE]
Count <- c(Count, Total=sum(Count))
Count
```

Save the total count as `Ans` for future computations.
```{r save_ans}
Ans <- Count["Total"]
```

The total number of pieces in the jar is `r Ans`.

## The Winners

We will consider a winner for each of the three groups.
To compute the winners, we first compute the error of each guess,
the difference between each guess and the answer, taking the absolute value.

```{r compute_error}
guesses <- guesses %>% mutate(Error = abs(Guess - Ans))
```


Winner has minimum `Error`:
```{r winner_by_group}
winners <- guesses %>% group_by(Group) %>%
  summarise(Person=Person[which.min(Error)],
    Guess=Guess[which.min(Error)], Error = min(Error), .groups="drop")
winners %>% knitr::kable()
```

## Wisdom of the Crowd

There is a theory that the average of all guesses for something like this
will be a good estimate of the true value.
We consider the average by Group as well as the overall average across all
three groups.
For the average we consider both the mean and the median.

Compute averages and their errors, make a new dataset with all guesses.

```{r average}
group_mean <- guesses %>% group_by(Group) %>%
  summarise(Guess=mean(Guess), n=n(), .groups="drop") %>%
  mutate(Person="Mean", Error=abs(Guess - Ans))
group_median <- guesses %>% group_by(Group) %>%
  summarise(Guess=median(Guess), n=n(), .groups="drop") %>%
  mutate(Person="Median", Error=abs(Guess - Ans)) %>%
  ungroup()
overall_mean <- guesses %>%
  summarise(Group="All", Person="AllMean", Guess=mean(Guess), n=n(),
    .groups="drop") %>%
  mutate(Error=abs(Guess - Ans))
overall_median <- guesses %>%
  summarise(Group="All", Person="AllMedian",
    Guess=median(Guess), n=n(), .groups="drop") %>%
  mutate(Error=abs(Guess - Ans))
crowd_est <- group_mean  %>% full_join(group_median,
    by=c("Group", "Guess", "n", "Person", "Error")) %>%
  full_join(overall_mean, by=c("Group", "Guess", "n", "Person", "Error")) %>%
  full_join(overall_median, by=c("Group", "Guess", "n", "Person", "Error"))
estimates <- group_mean %>% full_join(group_median,
    by=c("Group", "Guess", "n", "Person", "Error")) %>%
  full_join(overall_mean, by=c("Group", "Guess", "n", "Person", "Error")) %>%
  full_join(overall_median, by=c("Group", "Guess", "n", "Person", "Error")) %>%
  select(Group, Person, Guess, Error, n)
estimates
```

Combine with the individual winners:

```{r combine_estimates}
estimates <- estimates %>% full_join(winners,
    by=c("Group", "Person", "Guess", "Error")) %>%
  rename(Estimate=Person, Value=Guess) %>%
  arrange(Error)
estimates %>% knitr::kable()
```

## Visualizations

The distribution of all guesses
```{r hist_all, fig.width=9, fig.height=7}
guesses %>% ggplot(aes(x=Guess)) + geom_histogram(binwidth=10) +
  geom_vline(xintercept=Ans, color="red", size=0.5) +
  ggtitle("All Guess, Red Line Is the Answer")
```

The distribution of guesses by group:
```{r hist_group, fig.width=9, fig.height=7}
guesses %>% ggplot(aes(x=Guess)) + geom_histogram(binwidth=10) +
  facet_grid(~ Group) +
  geom_vline(xintercept=Ans, color="red", size=0.5)
  ggtitle("Guesses by Group, Red Line Is the Answer")
```

The crowd estimates along with all the guesses:
```{r estimates, fig.width=9, fig.height=7}
guesses %>% ggplot(aes(x=Group, y=Guess)) + geom_point(shape=1) +
  geom_hline(yintercept=Ans, color="red", size=0.15) +
  geom_point(data=estimates, aes(x=Group, y=Value, color=Estimate),
    size=3, shape=18) +
  ggtitle("Crowd Estimates, Winners and All Guesses (open circles), Red Line Is the Answer")
```

## Appendix
This is a reproducible document.
All code and data are available at https://github.com/spkaluzny/crowd_wisdom.

```{r sessionInfo}
sessionInfo()
```
