---
title: "Cork Counting and the Crowd"
author: "Stephen Kaluzny"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Challenge

Guess the number of corks in the plastic tub.
This was a typically activity at the Seattle TIBCO Office's Halloween Party.
There you could handle the physical jar of candy.
For this exercise, several photos of the tub were provided.

There is an idea that the Wisdom of the Crowd would provide a good
estimate of the count.
The story goes that guesses for a pig's weight at a state fair were highly
variable but the average of all the guesses was very close to the correct value.

The exercise was made available to
my extended family (Family).

```{r load_packages, echo=FALSE}
suppressPackageStartupMessages(library("here"))
suppressPackageStartupMessages(library("assertr"))
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("ggplot2"))
```

## The Data

```{r read_data}
guesses <- read.csv(here::here("data", "cork_guesses.csv")) %>% select(-Group)
```

Quick quality check on the data,
should be only one observations per `Person`.

```{r data_qc}
guesses %>% count(Person) %>% verify(n == 1) %>% success_logical()
```

There are `r nrow(guesses)` total observations (guesses).

The correct cork count is:
```{r save_ans}
(Ans <- 559)
```

## The Winner

To compute the winner, we first compute the error of each guess,
the difference between each guess and the answer, taking the absolute value.

```{r compute_error}
guesses <- guesses %>% mutate(Error = abs(Guess - Ans))
```

Winner has minimum `Error`:
```{r winner}
winners <- guesses %>%
  summarise(Person=Person[which.min(Error)],
    Guess=Guess[which.min(Error)], Error = min(Error))
winners %>% knitr::kable(format="html", table.attr = "style='width:30%;'")
```

## Wisdom of the Crowd

There is a theory that the average of all guesses for something like this
will be a good estimate of the true value.
For the average we consider both the mean and the median.

Compute averages and their errors, make a new dataset with all guesses.

```{r average}
overall_mean <- guesses %>%
  summarise(Person="Mean", Guess=mean(Guess)) %>%
  mutate(Error=abs(Guess - Ans))
overall_median <- guesses %>%
  summarise(Person="Median", Guess=median(Guess)) %>%
 mutate(Error=abs(Guess - Ans))
```

```{r estimates}
estimates <- guesses %>%
  full_join(overall_mean, by=c("Person", "Guess", "Error")) %>%
  full_join(overall_median, by=c("Person", "Guess", "Error")) %>%
  arrange(Error)
estimates %>% knitr::kable(format="html", table.attr = "style='width:30%;'")
```

## Visualizations

```{r meanmedianans}
estimates$Group <- rep("Guess", nrow(estimates))
estimates$Group[estimates$Person == "Mean"] <- "Mean"
estimates$Group[estimates$Person == "Median"] <- "Median"
estimates <- estimates %>%
  rbind(data.frame(Person="Answer", Guess=Ans, Error=0, Group="Answer"))
```

```{r, fig.width=3}
estimates %>%
  ggplot(aes(x="Corks", y=Guess, color=Group)) +
    geom_jitter(cex=2, height=0, width=0.10) +
    xlab("")
```

The distribution of all guesses.
```{r hist_all, fig.width=9, fig.height=7}
estimates %>% filter(Person != "Answer") %>%
  mutate(Person=factor(Person, levels=Person)) %>%
  ggplot(aes(x=Person, y=Guess)) +
    stat_summary(fun="mean", geom="bar", width=0.3) +
    geom_hline(yintercept=Ans, color="red", size=0.5) +
    coord_flip() +
    ggtitle("All Guesses, Red Line Is the Answer",
      subtitle="(ordered by error)") +
    theme_light()
```

We see from the graph that most people underestimated the count,
many were way too low.
Only two estimates, `susank` and `laurak` where above the actual cork count.
Since most of the estimates were too low, the crowd estimate based on the
median was not very good.
The estimate based on the mean did better
because of the very large over estimate by `laurak` which pulled the mean
closer to the true value.


## Appendix
This is a reproducible document.
All code and data are available at https://github.com/spkaluzny/crowd_wisdom.

```{r sessionInfo}
sessionInfo()
```
