---
title: "Cork Counting and the Crowd"
author: "Stephen Kaluzny"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Challenge

Guess the number of corks in the plastic tub.
This was a typically activity at the Seattle TIBCO Office's Halloween Party.
There you could handle the physical jar of candy.
For this exercise, several photos of the tub were provided.

There is an idea that the Wisdom of the Crowd would provide a good
estimate of the count.
The story goes that guesses for a pig's weight at a state fair were highly
variable but the average of all the guesses was very close to the correct value.

The exercise was made available to
my extended family (Family).

```{r load_packages, echo=FALSE}
suppressPackageStartupMessages(library("here"))
suppressPackageStartupMessages(library("assertr"))
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("ggplot2"))
```

## The Data

```{r read_data}
guesses <- read.csv(here::here("data", "cork_guesses.csv"))
```

Quick quality check on the data,
should be only one observations per `Person`.

```{r data_qc}
guesses %>% count(Person) %>% verify(n == 1) %>% success_logical()
```

There are `r nrow(guesses)` total observations (guesses).

The correct cork count is:
```{r save_ans}
(Ans <- 559)
```

## The Win

To compute the winner, we first compute the error of each guess,
the difference between each guess and the answer, taking the absolute value.

```{r compute_error}
guesses <- guesses %>% mutate(Error = abs(Guess - Ans))
```

Winner has minimum `Error`:
```{r winner}_by_group}
winners <- guesses %>%
  summarise(Person=Person[which.min(Error)],
    Guess=Guess[which.min(Error)], Error = min(Error))
winners %>% knitr::kable()
```

## Wisdom of the Crowd

There is a theory that the average of all guesses for something like this
will be a good estimate of the true value.
We consider the average by Group as well as the overall average across all
three groups.
For the average we consider both the mean and the median.

Compute averages and their errors, make a new dataset with all guesses.

```{r average}
overall_mean <- guesses %>%
  summarise(Group="Family", Person="Mean", Guess=mean(Guess), n=n()) %>%
  mutate(Error=abs(Guess - Ans))
overall_median <- guesses %>%
  summarise(Group="Family", Person="Median", Guess=median(Guess), n=n()) %>%
 mutate(Error=abs(Guess - Ans))
```

Combine with the individual winners:

```{r combine_estimates}
estimates <- estimates %>% full_join(winners,
    by=c("Group", "Person", "Guess", "Error")) %>%
  rename(Estimate=Person, Value=Guess) %>%
  arrange(Error)
estimates %>% knitr::kable()
```

## Visualizations

The distribution of all guesses
```{r hist_all, fig.width=9, fig.height=7}
guesses %>% ggplot(aes(x=Guess)) + geom_histogram(binwidth=10) +
  geom_vline(xintercept=Ans, color="red", size=0.5) +
  ggtitle("All Guess, Red Line Is the Answer")
```

## Appendix
This is a reproducible document.
All code and data are available at https://github.com/spkaluzny/crowd_wisdom.

```{r sessionInfo}
sessionInfo()
```
